\section{Related Work}
\label{sec:related_work}

The episodic operating model which is used for this thesis falls under the umbrella of Lifelong SLAM. The term lifelong SLAM has varied definitions throughout the literature, but for simplicity, this research adopts the definition provided by Shi et al. \cite{shiAreWeReady2020}, describing lifelong SLAM as the ability for a robot to generate, maintain, and localize within a map of a particular environment over an extended period of time. Unlike the standard SLAM operating model, lifelong SLAM operating models are expected to perform without the assumption of a static environment. This research falls under the topic of map maintenance, but has downstream effects on other operations related to lifelong SLAM.

Several research topics fall under lifelong SLAM, with those relevant to the work presented in this thesis discussed below. This research directly impacts the KV-SLAM map, so the downstream effects on other aspects of lifelong SLAM performance will be discussed at the end.

\subsection{Place Recognition over Long Time Horizons}

As previously discussed, place recognition refers to the ability to determine when an agent has entered a previously visited area. In KV-SLAM, this is commonly achieved using a bag-of-words approach over binary feature descriptors extracted from camera images \cite{camposORBSLAM3AccurateOpenSource2021}. However, these features can vary significantly over time due to changes in lighting, weather, scene layout, or viewpoint. Chen et al. \cite{chenDeepLearningFeatures2017} proposed using deep learning to identify features that are invariant to conditions and viewpoint, but they are still sensitive to geometric changes.

Milford et al. \cite{milfordSeqSLAMVisualRoutebased2012} improved robustness to time-of-day and weather variation by matching sequences of whole-image comparisons rather than relying on a single image or feature-based matches. However, this approach requires alignment between the image sequence and the reference trajectory, effectively requiring similar speed and direction of travel. Naseer et al. \cite{naseerVisionbasedMarkovLocalization2015} relaxed this constraint by computing a full probability distribution over possible locations, removing the need for trajectory alignment. Both of these approaches are sensitive to viewpoint variation, and can fail when the camera trajectory differs significantly from previously observed trajectories.

Learning-based methods like NetVLAD \cite{arandjelovicNetVLADCNNArchitecture2016} assign global descriptors to places rather than features, increasing robustness to viewpoint, conditions, and dynamic clutter. HF-Net \cite{sarlinCoarseFineRobust2019} extends this idea by utilizing global descriptors for large-scale place recognition and local learned features for refinement. The negative aspects of learning-based approaches tend to be their computational load, requiring significantly higher capability than methods that do not require a neural network.

While the methods discussed here do allow for place recognition to succeed despite significant visual and geometric changes, in general place recognition is most effective when the internal scene representation matches the true environment. To accomplish this, the system must be able to identify and correct discrepancies between the map and the current sensor readings.

\subsection{Change Detection for Map Maintenance}

Change detection in SLAM refers to identifying discrepancies between current sensor observations and the existing map. This functionality supports operations like relocalization and place recognition by acting as the mechanism through which outdated map data is identified so it can be updated. The implementation of change detection depends heavily on sensor modality and map representation.

Girardeau-Montaut et al. \cite{girardeau-montautCHANGEDETECTIONPOINTS2005} implemented point cloud change detection for laser scans of industrial sites using iterative closest point (ICP) \cite{arunLeastSquaresFittingTwo1987,beslMethodRegistration3D1992} to align scans, with Hausdorff distance \cite{aspertMESHMeasuringErrors2002} used to identify areas of change. Lin et al. \cite{linPointCloudChange2022} extended this to stereo visual SLAM in their system PPCA-VINS, which detects changes between a prior point cloud and current point measurements. The method of aligning point clouds and measuring disparity is effective, but requires both point clouds to be of similar scale and density to achieve optimal alignment. Additionally, these methods do not attempt to model global point observability, meaning features that still exist could be erroneously deleted.

Many modern methods utilize a learning-based approach to align and detect changes in dissimilar data types. Larsen et al. \cite{larsenChangeDetectionModel} % TODO: expand description of Larsen et al.

Lázaro et al. \cite{lazaroEfficientLongtermMapping2018} implemented map updates for ... % TODO: expand
Wilf et al. \cite{wilfCumulativeEvidenceScene2022} ... % TODO: expand
Qian et al. \cite{qianPOCDProbabilisticObjectLevel2022} ... % TODO: expand
Takeda et al. \cite{takedaLifelongChangeDetection2023} ... % TODO: expand
Schmid et al. \cite{schmidKhronosUnifiedApproach2024} ... % TODO: expand

\subsection{Dynamic Object Rejection}

Dynamic rejection is the process of identifying dynamic objects within a scene and rejecting image features that fall on them. This requires either tracking object motion to verify the object is dynamic, or performing image segmentation and semantic identification to categorize objects as either dynamic or static. It can also be accomplished without strict object identification using methods that detect movement in images (e.g., optical flow) or in point clouds. % TODO: add example/reference for point cloud motion methods

Saputra et al. \cite{saputraVisualSLAMStructure2019} ... % TODO: expand
Henein et al. \cite{heneinDynamicSLAMNeed2020} ... % TODO: expand
Mao et al. \cite{maoSemanticMaskingVisual2024} ... % TODO: expand

\subsection{Feature Persistence and Observability Modeling}

Feature persistence modeling refers to the estimation of whether a specific map feature will remain observable over time, producing per-feature metrics that can drive map maintenance decisions.

Derner et al. \cite{dernerChangeDetectionUsing2021} ... % TODO: expand
Berrio et al. \cite{berrioUpdatingVisibilityFeaturebased2019,berrioLongtermMapMaintenance2020} ... % TODO: expand
Alcantarilla et al. \cite{alcantarillaLearningVisibilityLandmarks2010} ... % TODO: expand
Hashemifar and Dantu \cite{hashemifarPracticalPersistenceReasoning2020} ... % TODO: expand
Rosen et al. \cite{rosenLifelongFeaturebasedMapping2016} ... % TODO: expand
Walcott-Bryant et al. \cite{walcott-bryantDynamicPoseGraph2012} ... % TODO: expand
