\subsection{Definitions and Functions}

Future sections explore the mathematical basis and implementation of the proposed system. To provide a common reference, useful definitions and constructs are defined here.

\newcommand{\viewpoint}{\ensuremath{v}}
\newcommand{\viewpointdomain}{\ensuremath{V}}
\newcommand{\dmin}{\ensuremath{d_{\text{min}}}}
\newcommand{\dmax}{\ensuremath{d_{\text{max}}}}
\newcommand{\observation}{\ensuremath{o}}
\newcommand{\observations}[1]{\ensuremath{\boldsymbol{O}^{#1}}}
\newcommand{\pe}{\ensuremath{P(E)}}
\newcommand{\estpe}{\ensuremath{\hat{P}(E)}}
\newcommand{\ps}{\ensuremath{P(S,v)}}
\newcommand{\psge}{\ensuremath{P(S,v|E)}}
\newcommand{\estpsge}{\ensuremath{\hat{P}(S,v|E)}}
\newcommand{\psgne}{\ensuremath{P(S,v|\lnot E)}}
\newcommand{\pegs}{\ensuremath{P(E|S,v)}}
\newcommand{\pegns}{\ensuremath{P(E|\lnot S,v)}}
\newcommand{\falsepos}{\ensuremath{\varepsilon_{\text{pos}}}}
\newcommand{\falseneg}{\ensuremath{\varepsilon_{\text{neg}}}}
\newcommand{\damping}{\ensuremath{\alpha}}
\newcommand{\observability}{\ensuremath{\rho}}
\newcommand{\observabilityupdate}{\ensuremath{\eta}}
\newcommand{\state}{\ensuremath{s}}
\newcommand{\feedback}{\ensuremath{\beta}}
\newcommand{\params}[1]{\ensuremath{\boldsymbol{\theta}_{#1}}}
\newcommand{\param}[2]{\ensuremath{\theta_{#1,#2}}}
\newcommand{\epe}{\ensuremath{\text{EPE}}}
\newcommand{\vbee}{\ensuremath{\text{VBEE}}}
\newcommand{\knn}{\ensuremath{\text{KNN}}}
\newcommand{\db}{\ensuremath{\text{DB}}}
\newcommand{\initialize}[1]{\ensuremath{\operatorname{Initialize}(#1)}}
\newcommand{\estimate}[1]{\ensuremath{\operatorname{Estimate}(#1)}}
\newcommand{\integrate}[1]{\ensuremath{\operatorname{Integrate}(#1)}}
\newcommand{\update}[1]{\ensuremath{\operatorname{Update}(#1)}}
\newcommand{\query}{\ensuremath{\operatorname{Query}()}}
\newcommand{\athreshold}{\ensuremath{\tau_{\text{angle}}}}
\newcommand{\feedbackthreshold}{\ensuremath{\tau_{\beta}}}
\newcommand{\existsthreshold}{\ensuremath{\tau_{E}}}

\paragraph{Definitions}
\defitem{$\dmin$ and $\dmax$}{The minimum and maximum distances at which a map point may be considered visible to the camera.}
\defitem{$\viewpoint$}{A viewpoint in the domain $\viewpoint \in \mathbb{S}^2 \times [\dmin, \dmax] \rightarrow \mathbb{R}^3$.}
\defitem{$\observation$}{An observation of a map point defined as $\observation = (\viewpoint, \state)$, where $\viewpoint$ is the viewpoint and $\state \in \{0.0, 1.0\}$ indicates whether the point was seen (1.0) or not seen (0.0) from $\viewpoint$.}
\defitem{$\observations{n}$}{A vector of $n$ observations in the form $\{\observation_0,\dots,\observation_n\}$.}
\defitem{$\pe$}{The probability that the map point exists.}
\defitem{$\ps$}{The probability that the map point is visible from a given viewpoint $\viewpoint$.}
\defitem{$\falseneg$}{The probability of a false negative (not observing a map point when it is actually observable).}
\defitem{$\falsepos$}{The probability of a false positive (observing a map point when it is actually unobservable).}
\defitem{$\damping$}{The damping coefficient applied to updates of $\pe$.}
\defitem{$\observability_t$}{The estimated overall observability of the map point at time $t$.}
\defitem{$\observabilityupdate$}{The damping coefficient applied to updates of $\observability$.}
\defitem{$\feedback$}{The feedback propagated from the Bayesian update step back to the observability model, calculated as $\estpe_t - \estpe_{t-1}$.}
\defitem{$\knn$}{An instance of the KNN observability model, a precursor to the Discrete Boundary model (see Section~\ref{sec:observability_models}).}
\defitem{$\db$}{An instance of the Discrete Boundary observability model.}
\defitem{$\epe$}{The Bayesian existence probability update step, now implemented internally within $\vbee$ rather than as a separate component.}
\defitem{$\vbee$}{An instance of the viewpoint-based existence estimator, which holds a reference to an observability model and performs Bayesian existence probability updates internally.}
\defitem{$\params{\knn}$}{A vector of parameters specific to the KNN model in the form $\{\param{\knn}{1},\dots,\param{\knn}{n}\}$, where $n$ is the number of parameters.}
\defitem{$\params{\db}$}{A vector of parameters for the Discrete Boundary model in the form $\{k, \text{ambiguous\_range}\}$, where $k$ is the ambiguity counter maximum and $\text{ambiguous\_range}$ controls the width of the ambiguous region's output range.}
\defitem{$\params{\epe}$}{The false negative and false positive rates $\{\falseneg,\falsepos\}$, now incorporated directly into $\params{\vbee}$.}
\defitem{$\params{\vbee}$}{A vector of parameters for the viewpoint-based existence estimator in the form $\{model,\estpe_0,\damping,\observability_0,\observabilityupdate,\falseneg,\falsepos\}$, where $model$ is the observability model instance.}

\paragraph{Functions}
\defitem{$\vbee.\initialize{\params{\vbee}}$}{Initializes the viewpoint-based existence estimator, including the observability model instance.}
\defitem{$\vbee.\update{\observation}$}{Updates the estimator with a new observation $\observation$, returning the updated existence probability $\estpe$.}
\defitem{$\vbee.\query$}{Queries the estimator for the current estimate of the map point's existence probability $\estpe$.}
\defitem{$\db.\initialize{\params{\db}}$}{Initializes the Discrete Boundary model with parameters $\params{\db}$.}
\defitem{$\db.\estimate{\viewpoint}$}{Returns the Discrete Boundary model's estimated probability of seeing the map point from viewpoint $\viewpoint$, given the point exists: $\estimate{\viewpoint} = \estpsge$.}
\defitem{$\db.\integrate{\observation}$}{Integrates a new observation $\observation$ into the Discrete Boundary model's internal bin state.}
\defitem{$\knn.\initialize{\params{\knn}}$}{Initializes the KNN model (precursor) with parameters $\params{\knn}$.}
\defitem{$\knn.\estimate{\viewpoint}$}{Returns the KNN model's estimated probability of seeing the map point from viewpoint $\viewpoint$: $\estimate{\viewpoint} = \estpsge$.}
\defitem{$\knn.\integrate{\observation, \feedback}$}{Integrates a new observation $\observation$ into the KNN model's stored observation buffer, using feedback $\feedback$ to determine whether the observation is incorporated.}
