\subsection{Contribution}

Through this research, we contribute three things to the subject of long-term map reuse in KV-SLAM. First, we release an open-source library which provides reusable access to the mathematical implementation of the VAMPE model. Next, a modified version of ORB-SLAM3 with the VAMPE model integrated is released for ease of further testing and research. Finally, we provide a collection of datasets covering situations particularly challenging for intermittently run KV-SLAM systems for demonstration and comparison of the system's capabilities.

% The library: a lightweight, viewpoint-aware model of map point existence, designed to be utilized in keypoint-based visual SLAM systems.

The VAMPE library is a C++ implementation of the viewpoint-aware probability shell, with hooks for simple integration into any existing KV-SLAM implementation. The library

% The SLAM implementation: An implementation of the library in ORB-SLAM 3 to demonstrate its effectiveness.

% The datasets: A set of datasets and experimental results demonstrating how the library improves SLAM performance.

Through this research, we introduce an incrementally updated directional confidence model for the existence of map points. This model differs from other point removal optimizations in several ways. First, this implementation avoids the use of neural networks, facilitating use on resource constrained hardware without facilities optimized to run them. Second, while other probability based point removal optimizations have been developed, this model introduces the idea of utilizing a continuously updated perspective dependent shell of metadata for each keypoint, which can be used to reduce the problem of point existence to a simple Bayesian update step. This implementation allows perspective of observation to play a role in the point's existence probability update step, and avoids some of the common a priori work such as prior estimation common to other point removal optimization techniques. This shell is implemented in both finite and continuous modalities, utilizing regular convex polyhedral shells in the finite case, and von-meiser fisher distributions on the sphere in the continuous case.

To facilitate future research, this model is released as an open-source library, which is compatible with any keypoint based visual SLAM implementation. Additionally, a collection of co-registered visual-inertial and LIDAR datasets is provided, containing instances of multiple traversals through the same environments with changes to scene contents. Information regarding the locations of these environmental changes is included in the dataset, facilitating the benchmarking of point removal optimization implementations.